A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-08-06 paper/elsarticle/elsarticle-template.tex
paper

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity). This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation. We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the target location in the scene plane and (ii) rotated to place the axes of the virtual projector in accordance with those of the camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system set up to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

@@

Calibrating a camera can be carried out by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ in pixels of the point~$\mathbf{X}_{i,j}$ to that image. The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}. The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern.

Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene \cite{moreno2012simple,zhang2009projector,zhang2006novel}. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors. A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.

@@

The hardware setup employed in the evaluation (cf.\ Figure \ref{fig:eval}(a)) comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information,Rupprecht2021}. The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset. In addition, we used a downwards-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view. The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.

@@

@@

We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, in a manner that enables intuitive placement the augmentations in accordance with the axes of an image of the scene acquired by a downwards-facing camera, and such that the desired dimensions of augmentations can be specified in metric terms. Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view. Our evaluation demonstrated our approach to produce compelling results at less time than the more cumbersome traditional manual approach to keystone correction.

A natural extension of this work would be to address the augmentation of non-planar scenes. To handle non-planar scenes would call for a change in how scene geometry is recovered and how warping of the image to be projected is carried out; the methodology we proposed for projector calibration could, however, be left unchanged.

>>>>

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity). This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation. We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the target location in the scene plane and (ii) rotated to place the axes of the virtual projector in accordance with those of the camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system set up to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

@@

A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses the point~$\mathbf{X}_{i,j}$ in the local coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ in pixels of $\mathbf{X}_{i,j}$ to that image. The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}. Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene. Since our approach to projector calibration relies on a calibrated camera, what remains of this section is concerned primarily with the recovery of 2D-3D correspondences first in support of calibrating cameras, then of projectors. A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.

@@

The hardware setup employed in the evaluation (cf.\ Figure \ref{fig:eval}(a)) comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information,Rupprecht2021}. The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset. In addition, we used a downwards-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view. The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m. At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was ca.\ 3.5 mm.

@@

... At height above ground of 3.5 m, the pixel size was approximately 3.5 mm. Standard deviation of all measured sides was in all instances below the pixel size.

@@

We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically as a function of the geometry relating projector and scene plane, exploiting a steerable mirror system to enable augmentation exceeding the bounds of the projector's own immediate field of view. Our method produces this effect in a manner that enables intuitive placement of the augmentations in accordance with the $X$- and $Y$-axes of a camera placed to face downwards towards the scene plane. Moreover, our method allows for specifying the desired dimensions of augmentations in metric terms, thus enabling consistent scaling. Our evaluation demonstrated our approach to produce compelling results---with mean and standard deviation of errors in measured lengths within the downwards-facing camera's pixel size---at less time than a more cumbersome manual approach to keystone correction.

A natural extension of this work would be to consider the impact of varying calibration patterns, their dimensions, and their scale relative to the camera. Still another would be to address the augmentation of non-planar scenes. To handle non-planar scenes would call for a change in how scene geometry is recovered and how warping of the image to be projected is carried out; the methodology we proposed for projector calibration could, however, be left unchanged.
